Symmetric Peer-to-Peer Applications

### 2022-08-21 tpd-22.tex
...

\IEEEtitleabstractindextext{%
\begin{abstract}
In multi-node collaborative applications, such as shared online documents,
users can interact remotely and yet share the exact same experience as if they
were in a single machine.
%
In this work, we propose a middleware for \emph{symmetric peer-to-peer
applications} in which decentralized instances can broadcast asynchronous
events and yet conform to identical behavior.
%
Nodes are allowed to join and leave the network at any time.
Application developers must adhere to a restricted API, which is
deterministic, stateless, and only supports pre-allocated memory.
%
The middleware is responsible for synchronizing the events in a global shared
timeline across the network.
Based on memory snapshots and deterministic simulation, a ``time machine'' can
rollback conflicting nodes to resynchronize their state.
%
TODO: application, results
\end{abstract}

@@

In this work, we propose \emph{symmetric peer-to-peer applications}, which
do not rely on central servers for coordination.
Nodes in the application form a dynamic graph network and communicate only
with direct neighbours.
Events are flooded in the graph and are triggered locally with a small delay
to consider the network latency.
To deal with events received out of order or too late, a distributed time
machine can rollback nodes to a previous state and reapply events in order and
in time.

@@

\begin{figure*}[t]
  \centering
  \includegraphics[width=\linewidth]{algos}
  \caption{
    \label{fig.algos}
  }
\end{figure*}

@@

An antagonistic approach to deterministic reactions to events is to model the
application with conflict-free replicated data types (CRDTs)~\cite{crdts}.
CRDTs are designed in such a way that all operations are commutative, so that
the order in which they are applied does not affect the final outcome.
%
As detailed in Figure~\ref{fig.algos}, peers can communicate operations
directly to each other with no central authority.
Also, since operations need not to be ordered, they can be applied at the
very first moment they are generated or received, even if the peers are
offline.
%
Since operations must be commutative, it is not possible to timestamp them
in a global shared clock.
Hence, there is no notion of a timeline in which peers go through
bit-identical steps, but they do eventually reach the same state.
%
On the one hand, the main advantage of CRDTs is the support for local-first
software~\cite{local}, since they can work offline in the same way as online.
On the other hand, they provide very restrictive APIs (the CRDTs themselves),
and do not support real-time consensus among peers.
%
As indicated in Figure~\ref{fig.table}, CRDTs work on decentralized networks,
in which nodes advance independently, and in which operations can be applied
immediately, even while offline.
%
Automerge along with its accompanying Hypermerge peer-to-peer infrastructure
is an example of an implementation of decentralized
CRDTs~\cite{p2p.automerge,p2p.pushpin}.

@@

\section{The API \& Middleware}
\label{sec.tml}

In this section, we describe the architecture of our middleware and also the
API in which developers need to conform.
%
We guide our discussion through a didactic example in which multiple users use
the arrow keys to change the direction of a moving rectangle.

>>>>

\IEEEtitleabstractindextext{%
\begin{abstract}
In multi-node collaborative applications, such as shared online documents,
users can interact remotely and yet share the exact same experience as if they
were in a single machine.
%
In this work, we propose a middleware for \emph{symmetric peer-to-peer
applications} in which decentralized instances can broadcast asynchronous
events and yet conform to identical behavior.
%
Nodes are allowed to join and leave the network at any time.
Application developers must adhere to a restricted API, which is
deterministic, stateless, and only supports pre-allocated memory.
%
The middleware is responsible for synchronizing the events in a global shared
timeline across the network.
Based on memory snapshots and deterministic simulation, a ``time machine'' can
rollback conflicting peers to resynchronize their state.
%
TODO: application, results
\end{abstract}

@@

In this work, we propose \emph{symmetric peer-to-peer applications}, which
do not rely on central servers for coordination.
Nodes in the application form a dynamic graph network and communicate only
with direct neighbours.
Events are flooded in the graph and are triggered locally with a small delay
to consider the network latency.
To deal with events received out of order or too late, a distributed time
machine can rollback peers to a previous state and reapply events in order and
in time.

@@

\begin{figure*}[t]
  \centering
  \includegraphics[width=\linewidth]{algos}
  \caption{
    \label{fig.algos}
    Four approaches to symmetric distributed applications: Croquet, GALS,
    CRDTs, and this work.
    The vertical arrows in parallel represent the timelines in nodes in the
    network.
    The arrows crossing nodes represent communication between them.
  }
\end{figure*}

@@

An antagonistic approach to deterministic reactions to events is to model the
application with conflict-free replicated data types (CRDTs)~\cite{crdts}.
CRDTs are designed in such a way that all operations are commutative, so that
the order in which they are applied does not affect the final outcome.
%
As detailed in Figure~\ref{fig.algos}, peers can communicate operations
directly to each other with no central authority.
Also, since operations need not to be ordered, they can be applied at the
very first moment they are generated or received, even if the peers are
offline.
%
Since operations must be commutative, it is not possible to timestamp them
in a global shared clock.
Hence, there is no notion of a timeline in which peers go through
bit-identical steps, but they do eventually reach the same state.
%
On the one hand, the main advantage of CRDTs is the support for local-first
software~\cite{local}, since they can work offline in the same way as online.
On the other hand, they provide very restrictive APIs (the CRDTs themselves),
and do not support real-time consensus among peers.
%
As indicated in Figure~\ref{fig.table}, CRDTs work on decentralized networks,
in which peers advance independently, and in which operations can be applied
immediately, even while offline.
%
Automerge along with its accompanying Hypermerge peer-to-peer infrastructure
is an example of an implementation of decentralized
CRDTs~\cite{p2p.automerge,p2p.pushpin}.

@@

\section{The Middleware \& Programming API}
\label{sec.tml}

In this section, we describe the architecture of our middleware and also the
API in which developers have to conform with.

We guide our discussion through the didactic example of Figure~\ref{fig.move},
in which multiple users use the arrow keys to control the direction of a
moving rectangle.
The six instances are connected in a peer-to-peer network, such that some
peers can only communicate indirectly.
%
The application is symmetric in the sense that if any peer presses a key, an
event is broadcast and applied in all peers simultaneously, as if they were a
single mirrored instance.
At any moment, if a peer pauses the application, all peers are guaranteed to
pause exactly at the same frame with the rectangle at the same position.
%
A video

\begin{figure}[t]
    \centering
    \includegraphics[width=\linewidth]{move}
    \caption[XXX] {
	A moving rectangle is controlled by peers connected indirectly.
	Users can press \code{↑ ↓ → ←} to change the direction of the
	rectangle or \code{SPACE} to pause the application.
	The inputs are applied simultaneously in all peers, as if they were
	mirrors of a single application.
        Video: \url{http://youtube.com/TODO}
        \label{fig.move}
    }
\end{figure}
